---
title: "covid 19"
author: "Jaelynn Schultz"
date: "Summer 2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex


Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

According to this chart, more men than women have died from this virus with 44 men dead compared to only 25 women. In addition to this, there appear to be more women in the study, as there are a total of 1856 women and only 1455 men. This would insinuate that considerably more men are dying from this virus in comparison to women. However, more testing needs to be done to confirm if this difference is statistically significant.

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

`

  





  
## Numerical results


## Inferential results 




## Conclusions

